---
title: "Assignment9"
author: "Kota"
date: "12/2/2016"
output: pdf_document
---
```{r}
cdcbeacpi <- read.table(file="cdcbeacpi.txt", stringsAsFactors = F)
```
```{r}
library(ggplot2)
```

# problem 4: A graph of the inflation-adjusted price of cigarettes by state 1970–2014
```{r}
# rprice can be racpp
p4 <- cdcbeacpi[,c("year", "name", "fstprp", "fstpp", "inflation_factor")]
p4$price <- p4$fstpp*100/p4$fstprp
p4$rprice <- p4$price * p4$inflation_factor
g1 <- ggplot(data = p4, aes(x = year, y = rprice, group = name, color = name))
g1 + geom_line() + theme(legend.position = "bottom")
# I found rprice can be racpp after calculating
```

# Problem 5: A graph of average cigarette consumption by state 1970–2014
```{r}
g2 <- ggplot(data = cdcbeacpi, aes(x = year, y = cc, group = name, color = name))
g2 + geom_line() + theme(legend.position = "bottom")
```

# Problem 6: A graph of average cigarette consumption by year (across all states) 1970-2014
```{r}
p6 <- aggregate(cc ~ year, FUN = mean, data = cdcbeacpi)
g3 <- ggplot(data = p6, aes(x = year, y = cc))
g3 + geom_line()
```

# Problem 7: A graph of average price by year (across all states) 1970–2014
```{r}
p7 <- aggregate(racpp ~ year, FUN = mean, data = cdcbeacpi)
g4 <- ggplot(data = p7, aes(x = year, y = racpp))
g4 + geom_line()
```

# problem 8: Scatterplot of consumption (Y axis) and inflation-adjusted prices (X axis) 
```{r}
rm("mnd")
fit1 <- lm(cc ~ racpp, data = cdcbeacpi)
summary(fit1)

mnd <- data.frame(racpp = seq(1, 10.5, 0.001))
mnd$pred <- predict(fit1, mnd)
```

# square
```{r}
fit2 <- lm(cc ~ racpp + I(racpp^2), data = cdcbeacpi)
summary(fit2)

mnd$pred_square <- predict(fit2, mnd)
```

# cubic
```{r}
fit3 <- lm(cc ~ racpp + I(racpp^2) + I(racpp^3), data = cdcbeacpi)
summary(fit3)

mnd$pred_cubic <- predict(fit3, mnd)

plot(x = cdcbeacpi$racpp, y = cdcbeacpi$cc, pch = 20, col = alpha("black", 0.4))
lines(pred ~ racpp, col = "red", data = mnd)
lines(pred_square ~ racpp, col = "blue", data = mnd)
lines(pred_cubic ~ racpp, col = "green", data = mnd)
legend("topright", c("linear", "square", "cube"), lty = c(1,1,1),
       col = c("red","blue","green"))
rm("mnd")
```

Adjusted R-square: liner: 0.5729  quad: 0.6097  cubic: 0.6188
The relationship between cc and racpp doesn’t look liner, and data explains liner cubic model explains it better. 

# problem 9: Repeat the scatterplot but now using log consumption and log prices 
```{r}
log_racpp <- log(cdcbeacpi$racpp)
log_cc <- log(cdcbeacpi$cc)
p9 <- data.frame(log_racpp, log_cc)
```

```{r}
fit4 <- lm(log_cc ~ log_racpp, data = p9)
summary(fit4)

mnd <- data.frame(log_racpp = seq(0.2, 2.5, 0.001))
mnd$pred <- predict(fit4, mnd)
```

# square
```{r}
fit5 <- lm(log_cc ~ log_racpp + I(log_racpp^2), data = p9)
summary(fit5)

mnd$pred_square <- predict(fit5, mnd)
```

# cubic
```{r}
fit6 <- lm(log_cc ~ log_racpp + I(log_racpp^2) + I(log_racpp^3), data = p9)
summary(fit6)

mnd$pred_cubic <- predict(fit6, mnd)

plot(x = log_racpp, y = log_cc, pch = 20, col = alpha("black", 0.4))
lines(pred ~ log_racpp, col = "red", data = mnd)
lines(pred_square ~ log_racpp, col = "blue", data = mnd)
lines(pred_cubic ~ log_racpp, col = "green", data = mnd)
legend("topright", c("linear", "square", "cube"), lty=c(1,1,1),
       col=c("red","blue","green"))
```

Adjusted R-square: liner: 0.6897  quad: 0.6996  cubic: 0.7049
The relationship between cc and racpp looks liner, and data explains liner cubic model explains it a little bit better. 

# Problem 3: 3.	A table that shows your variable names and their definitions.

Documentation for cdcbeacpi Data

The datafile cdcbecpi contains data on all states about Cigarette in the US from 1970 to 2014. These data were provided by Centers for Disease Control and Prevention, which annually tracks data concerning about Tobacco including consumption of cigarette and governments’ policy by states. Bureau of Economic Analysis supplied data of income such as personal income and per capita income by states, and consumer price index is from Bereau of Labor Statistics. It has 2295 observations of 14 variables.
