---
title: "Step2: process and analyze data"
author: "Kota"
date: "12/11/2016"
---
```{r}
download.file("https://raw.githubusercontent.com/wscrdzg/econometrics/master/combined_data.txt",
              destfile = "combined_data.txt")
combined <- read.table(file="combined_data.txt", stringsAsFactors = F)
```

```{r}
library(ggplot2)
library(tidyr)
library(AER)
```

# Price 
A graph of the inflation-adjusted price of cigarettes by state 1970–2014
calculate the price use racpp + rgstpp
```{r}
combined$ravgprs <- combined$racpp + combined$rgstpp

g1 <- ggplot(data = combined, aes(x = year, y = ravgprs, group = name, color = name))
g1 + geom_line() + theme(legend.position = "none")
```

# Consumption by state
A graph of average cigarette consumption by state 1970–2014
```{r}
g2 <- ggplot(data = combined, aes(x = year, y = cc, group = name, color = name))
g2 + geom_line() + theme(legend.position = "none")
```

# Consumption by year
A graph of average cigarette consumption by year (across all states) 1970-2014
```{r}
p6 <- aggregate(cc ~ year, FUN = mean, data = combined)
g3 <- ggplot(data = p6, aes(x = year, y = cc))
g3 + geom_line()
```
# Average price by year
A graph of average price by year (across all states) 1970–2014
```{r}
p7 <- aggregate(ravgprs ~ year, FUN = mean, data = combined)
g4 <- ggplot(data = p7, aes(x = year, y = ravgprs))
g4 + geom_line()
```

# Relation of consumption and price

linear
```{r}
fit1 <- lm(cc ~ ravgprs, data = combined)
summary(fit1)

mnd <- data.frame(ravgprs = seq(1, 10.5, 0.001))
mnd$pred <- predict(fit1, mnd)
```

square
```{r}
fit2 <- lm(cc ~ ravgprs + I(ravgprs^2), data = combined)
summary(fit2)

mnd$pred_square <- predict(fit2, mnd)
```

cubic
```{r}
fit3 <- lm(cc ~ ravgprs + I(ravgprs^2) + I(ravgprs^3), data = combined)
summary(fit3)

mnd$pred_cubic <- predict(fit3, mnd)
```
Scatterplot of consumption (Y axis) and inflation-adjusted prices (X axis)
```{r}
plot(x = combined$ravgprs, y = combined$cc, pch = 20, col = alpha("black", 0.4))
lines(pred ~ ravgprs, col = "red", data = mnd)
lines(pred_square ~ ravgprs, col = "blue", data = mnd)
lines(pred_cubic ~ ravgprs, col = "green", data = mnd)
legend("topright", c("linear", "square", "cube"), lty = c(1,1,1),
       col = c("red","blue","green"))
rm("mnd")
```

Adjusted R-square: liner: 0.5814 quad: 0.6172 cubic: 0.6258 The relationship between cc and racpp doesn’t look liner, and data explains liner cubic model explains it better.

# Relation: ln(consumption) and ln(price)

```{r}
log_ravgprs <- log(combined$ravgprs)
log_cc <- log(combined$cc)
p9 <- data.frame(log_ravgprs, log_cc)
```

linear
```{r}
fit4 <- lm(log_cc ~ log_ravgprs, data = p9)
summary(fit4)

mnd <- data.frame(log_ravgprs = seq(0.2, 2.5, 0.001))
mnd$pred <- predict(fit4, mnd)
```

square
```{r}
fit5 <- lm(log_cc ~ log_ravgprs + I(log_ravgprs^2), data = p9)
summary(fit5)

mnd$pred_square <- predict(fit5, mnd)
```

cubic
```{r}
fit6 <- lm(log_cc ~ log_ravgprs + I(log_ravgprs^2) + I(log_ravgprs^3), data = p9)
summary(fit6)

mnd$pred_cubic <- predict(fit6, mnd)
```
Repeat the scatterplot but now using log consumption and log prices
```{r}
plot(x = log_ravgprs, y = log_cc, pch = 20, col = alpha("black", 0.4))
lines(pred ~ log_ravgprs, col = "red", data = mnd)
lines(pred_square ~ log_ravgprs, col = "blue", data = mnd)
lines(pred_cubic ~ log_ravgprs, col = "green", data = mnd)
legend("topright", c("linear", "square", "cube"), lty=c(1,1,1),
       col=c("red","blue","green"))
```

Adjusted R-square: liner: 0.6952 quad: 0.7067 cubic: 0.7127 The relationship between cc and racpp looks liner, and data explains liner cubic model explains it a little bit better.

# ln(cc) ~  |  Instrument variable: sales tax
Use sales tax as an instrument for price
```{r}
fit.iv1 <- ivreg(log(cc) ~ log(racpp) | rgstpp, data = combined)
summary(fit.iv1)
coeftest(fit.iv1, vcov = vcovHC(fit.iv1, "HC1"))
```

Use sales tax as an instrument for price
```{r}
fit.iv2 <- ivreg(log(cc) ~ log(racpp) + log(rpcpi) | rgstpp + log(rpcpi), data = combined)
summary(fit.iv2)
coeftest(fit.iv2, vcov = vcovHC(fit.iv2, "HC1"))
```

Use cigarette-specific tax as an instrument for price
```{r}
fit.iv3 <- ivreg(log(cc) ~ log(racpp) + log(rpcpi) | rstpp + log(rpcpi), data = combined)
summary(fit.iv3)
coeftest(fit.iv3, vcov = vcovHC(fit.iv3, "HC1"))
```

# delta ln()
now use change in log as regressor
```{r}
data96 <- subset(combined, year == 1996)
data14 <- subset(combined, year == 2014)
data96_14 <- merge(data96, data14, by = "name", suffixes = c("96", "14"))

data96_14$cc96_log <- log(data96_14$cc96)
data96_14$cc14_log <- log(data96_14$cc14)
data96_14$cc_dif_log <- data96_14$cc14_log - data96_14$cc96_log

data96_14$ravgprs96_log <- log(data96_14$ravgprs96)
data96_14$ravgprs14_log <- log(data96_14$ravgprs14)
data96_14$ravgprs_dif_log <- data96_14$ravgprs14_log - data96_14$ravgprs96_log

data96_14$rpcpi96_log <- log(data96_14$rpcpi96)
data96_14$rpcpi14_log <- log(data96_14$rpcpi14)
data96_14$rpcpi_dif_log <- data96_14$rpcpi14_log - data96_14$rpcpi96_log

data96_14$rgstpp_dif <- data96_14$rgstpp14 - data96_14$rgstpp96 # gross sales tax

data96_14$rgctr_dif <- data96_14$rgctr14 - data96_14$rgctr96 # gross specific tax
```

Use sales tax as an instrument for price
```{r}
fit.iv4 <- ivreg(cc_dif_log ~ ravgprs_dif_log + rpcpi_dif_log | 
                  rgstpp_dif + rpcpi_dif_log, data = data96_14)
summary(fit.iv4)
coeftest(fit.iv4, vcov = vcovHC(fit.iv4, "HC1"))
```

Same as above but use the cigarette-specific tax as an instrument for price
```{r}
fit.iv5 <- ivreg(cc_dif_log ~ ravgprs_dif_log + rpcpi_dif_log | 
                  rgctr_dif + rpcpi_dif_log, data = data96_14)
summary(fit.iv5)
coeftest(fit.iv5, vcov = vcovHC(fit.iv5, "HC1"))
```

Same as above but both the sales tax and the cigarette-specific tax as instruments for price
```{r}
fit.iv6 <- ivreg(cc_dif_log ~ ravgprs_dif_log + rpcpi_dif_log | 
                  rgstpp_dif + rgctr_dif + rpcpi_dif_log, data = data96_14)
summary(fit.iv6)
coeftest(fit.iv6, vcov = vcovHC(fit.iv6, "HC1"))
